Whittaker Models for Real Groups

### 2024-05-22 whittaker.tex
Jolted down the right place in Rossmann, to be completed
Fixed SL(2) formulas; notes in an appendix

Part~(c) follows from~\cite{rossmann_limit_orbits}. (Where?)

@@

\subsubsection{The Sekiguchi correspondence for principal nilpotent orbits}\label{sec:concrete_sek}
\footnote{AA 2024-IV-15: should correct the formulas here.}
Let us spell out the special case of the Sekiguchi correspondence which is of interest here. We follow~\cite[Section~2]{AVAV}.

Suppose $E_\R\in \Op\cap i\g(\R)$. Then we may choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ with $F_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$.
The Sekiguchi correspondence takes the $G(\R)$-orbit of~$E_\R$ to the $K$-orbit of
$$
E_\theta=\frac12(-iE_\R-iF_\R+H_\R)\in \N\cap \s.
$$

Computing the map in the other direction goes as follows.
Suppose $E_\theta\in \Op\cap\s$. Choose an $SL(2)$-triple  $(E_\theta H_\theta,F_\theta)$ with
$F_\theta\in\s$, which implies $H_\theta\in\k$. After conjugating by $K$ we may further assume $\sigma(E_\theta)=F_\theta$.
Then the Sekiguchi correspondence takes the $K$-orbit of $E_\theta$ to the $G(\R)$-orbit of 
\begin{equation}\label{def_E_R}
E_\R=\frac i2(E_\theta-F_\theta-H_\theta)\in\N\cap i\g(\R).
\end{equation}

@@

@@

\end{document}
%%% Local Variables: 
%%% mode: latex
%%% TeX-master: t
%%% End: s_{1}:0 -> -1 6

>>>>

Part~(c) follows from~\cite{rossmann_limit_orbits}. Update May 1: it seems the right place is Lemma B, and to go from that to the statement is a matter of chasing definitions and using the asymptotic cone. I (=AA) will have a look.

@@

\subsubsection{The Sekiguchi correspondence for principal nilpotent orbits}\label{sec:concrete_sek}
\footnote{JA These formulas should be right now, see the Appendix}
Let us spell out the special case of the Sekiguchi correspondence which is of interest here. We follow~\cite[Section~2]{AVAV}.

Suppose $E_\R\in \Op\cap i\g(\R)$. Then we may choose an $SL(2)$-triple $(H_\R,E_\R,F_\R)$ with $F_\R$  contained in $i\g(\R)$,
which implies $H_\R\in \g(\R)$. 
After conjugating by~$G(\R)$ we may further assume $\theta(E_\R)=-F_\R$.
The Sekiguchi correspondence takes the $G(\R)$-orbit of~$E_\R$ to the $K$-orbit of
$$
E_\theta=\frac12(E_\R-F_\R+H_\R)\in \N\cap \s.
$$

Computing the map in the other direction goes as follows.
Suppose $E_\theta\in \Op\cap\s$. Choose an $SL(2)$-triple  $(E_\theta, H_\theta,F_\theta)$ with
$F_\theta\in\s$, which implies $H_\theta\in\k$. After conjugating by $K$ we may further assume $\sigma(E_\theta)=F_\theta$.
Then the Sekiguchi correspondence takes the $K$-orbit of $E_\theta$ to the $G(\R)$-orbit of 
\begin{equation}\label{def_E_R}
E_\R=\frac 12(E_\theta-F_\theta-H_\theta)\in\N\cap i\g(\R).
\end{equation}

@@

\subsection{Appendix: $SL(2,\R)$ calculation}

This is not for publication, just to make sure we have the formulas in Section \ref{sec:concrete_sek} correct.
These are:
$$
E_\theta=\frac12(E_\R-F_\R+H_\R)\in \N\cap \s.
$$
and
$$
E_\R=\frac 12(E_\theta-F_\theta-H_\theta)\in\N\cap i\g(\R).
$$

Recall $\sigma(X)=\overline X$ and $\theta(X)=-X^t$.

Take
$$
E_\R=\begin{pmatrix}0&i\\0&0\end{pmatrix},
F_\R=\begin{pmatrix}0&0\\-i&0\end{pmatrix},
H_\R=\begin{pmatrix}1&0\\0&-1\end{pmatrix}
$$
These satisfy the standard $SL(2)$ commutation relations, and $E_\R,F_\R$ are in the $-1$-eigenspace of $\sigma$,
and $\theta(E_\R)=F_\R, \theta(H_\R)=-H_\R$.

Set $g=\frac 1{\sqrt 2}\begin{pmatrix}1&i\\i&1
\end{pmatrix}$, and take $E_\theta,F_\theta,H_\theta$ be the conjugates of $E_\R,F_\R,H_\R$ by $g$:
$$
E_\theta=\frac12\begin{pmatrix}1&i\\i&-1\end{pmatrix},
F_\theta=\frac12\begin{pmatrix}1&-i\\-i&-1\end{pmatrix},
H_\theta=\begin{pmatrix}0&-i\\i&0\end{pmatrix},
$$

These satisfy the standard $SL(2)$ commutation relations, and $E_\theta,F_\theta$ are in the $-1$-eigenspace of $\theta$.
Also $\sigma(E_\theta)=F_\theta,\sigma(H_\theta)=-H_\theta$.

Furthermore:
$$
\begin{aligned}
  E_\theta&=\frac12(E_\R-F_\R+H_\R)\\
  F_\theta&=\frac12(-E_\R+F_\R+H_\R)\\
  H_\theta&=-E_\R-F_\R
\end{aligned}
$$
and
$$
\begin{aligned}
  E_\R&=\frac12(E_\theta-F_\theta-H_\theta)\\
  F_\R&=\frac12(-E_\theta+F_\theta-H_\theta)\\
  H_\R&=E_\theta+F_\theta
\end{aligned}
$$

@@

\end{document}
